---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
subtitle: "Assignment 2 - Due date 01/26/22"
author: "Yu Hai"
output:
     pdf_document:
         latex_engine: xelatex
geometry: margin=2.54cm 
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

## Submission Instructions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github.   

Once you have the file open on your local machine the first thing you will do is change "Student Name" on line 4 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A02_Sp22.Rmd"). Submit this pdf using Sakai. 

## R packages

R packages needed for this assignment:"forecast","tseries", and "dplyr". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
library(forecast)
library(tseries)
library(dplyr)
library(readxl)
library(knitr)
```
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60), tidy=TRUE) 
```


## Data set information  

Consider the data provided in the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx" on our **Data** folder. The data comes from the US Energy Information and Administration and corresponds to the January 2022 Monthly Energy Review. The spreadsheet is ready to be used. Use the command $read.table()$ to import the data in R or $panda.read\_excel()$ in Python (note that you will need to import pandas package). }

```{r}
getwd()
#Importing data set
raw_RE_data<-read_excel(path="C:/Users/lenovo/Desktop/Spring_2022/ENV790/ENV790_TimeSeriesAnalysis_Sp2022/Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",skip=10)
raw_RE_data
```


## Question 1
You will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. Use the command head() to verify your data.

```{r}
sub_RE_data <- raw_RE_data[-c(1),4:6]
head(sub_RE_data)
```


## Question 2
Transform your data frame in a time series object and specify the starting point and frequency of the time series using the function ts().

```{r}
RE_data <- cbind(raw_RE_data[-c(1),1],sub_RE_data[,])
ts_RE_data <- ts(RE_data[,2:4], start=c(1973, 1), end=c(2021, 09), frequency=12)
head(ts_RE_data)
```



## Question 3
Compute mean and standard deviation for these three series.

```{r}
RE_data$`Total Biomass Energy Production`<-as.numeric(RE_data$`Total Biomass Energy Production`)
RE_data$`Total Renewable Energy Production`<-as.numeric(RE_data$`Total Renewable Energy Production`)
RE_data$`Hydroelectric Power Consumption`<-as.numeric(RE_data$`Hydroelectric Power Consumption`)
mean(RE_data$`Total Biomass Energy Production`)
mean(RE_data$`Total Renewable Energy Production`)
mean(RE_data$`Hydroelectric Power Consumption`)
sd(RE_data$`Total Biomass Energy Production`)
sd(RE_data$`Total Renewable Energy Production`)
sd(RE_data$`Hydroelectric Power Consumption`)
```


## Question 4
Display and interpret the time series plot for each of these variables. Try to make your plot as informative as possible by writing titles, labels, etc. For each plot add a horizontal line at the mean of each series in a different color.

```{r}
library(ggplot2)
ggplot(RE_data, aes(x=Month)) +
            geom_line(aes(y=RE_data$`Total Biomass Energy Production`)) +xlab("Year") + ylab("Energy Production (Trillion Btu)") + labs(title="Total Biomass Energy Production from 1973 to 2021")+geom_hline(yintercept = mean(RE_data$`Total Biomass Energy Production`), color="red")+theme_minimal()

ggplot(RE_data, aes(x=Month)) +
            geom_line(aes(y=RE_data$`Total Renewable Energy Production`)) +xlab("Year") + ylab("Energy Production (Trillion Btu)")+labs(title="Total Renewable Energy Production from 1973 to 2021")+geom_hline(yintercept = mean(RE_data$`Total Renewable Energy Production`), color="red")+theme_minimal() 

ggplot(RE_data, aes(x=Month)) +
            geom_line(aes(y=RE_data$`Hydroelectric Power Consumption`)) +xlab("Year") + ylab("Power Consumption (Trillion Btu)")+labs(title="Hydroelectric Power Consumption from 1973 to 2021")+geom_hline(yintercept = mean(RE_data$`Hydroelectric Power Consumption`), color="red")+theme_minimal()

```
Total biomass energy production shows a overall increasing trend. The most significant increase is from 2000 to 2010, while between 1990 and 2000 the increasing trend is weak but the variation within each year is large.

Total renewable energy production shows an overall increasing trend. This trend is not so obvious before 2000 and becomes significant since 2000. The seasonality is significant and relatively constant from 1973 to 2021. 

The hydroelectric power consumption doesn't show a clear increasing or decreasing trend over the years of observation, though there is always a great fluctuation between observations within each year. 

## Question 5
Compute the correlation between these three series. Are they significantly correlated? Explain your answer.

```{r}
cor(RE_data[,2:4])
```
According to the correlation coefficients, total biomass energy production shows a strong positive correlation with renewable energy production (0.92), while it shows a weak negative correlation with hydroelectric power consumption(-0.28). The renewable energy production also shows a weak negative correlation with hydroelectric power consumption(-0.0057).


## Question 6
Compute the autocorrelation function from lag 1 up to lag 40 for these three variables. What can you say about these plots? Do the three of them have the same behavior?

```{r}
Biomass_acf=Acf(ts_RE_data[,1],lag.max=40, type="correlation", plot=TRUE)
Renewable_acf=Acf(ts_RE_data[,2],lag.max=40, type="correlation", plot=TRUE)
Hydro_acf=Acf(ts_RE_data[,3],lag.max=40, type="correlation", plot=TRUE)

```
Biomass; ACF at all lags from 1 to 40 are positive (i.e. the correlation between Y1 and Y2,Y2...,Y40 are all positive), and there is a weak seasonality observed.

Renewable: Similar as the graph for Biomass energy production, ACFs at all lags are positive, and there is a stronger seasonality observed.

Hydroelectric: There are both positive and negative ACFs and there is a strong seasonality observed.

The three ACF graphs show different behaviors, but the common thing is that the absolute values of ACFs become smaller as the lag time goes up, and this is because the autocorrelation is weaker between the variables that are further away in time. 

## Question 7
Compute the partial autocorrelation function from lag 1 to lag 40 for these three variables. How these plots differ from the ones in Q6?

```{r}
Biomass_pacf=Pacf(ts_RE_data[,1],lag.max=40, plot=TRUE)
Renewable_pacf=Pacf(ts_RE_data[,2],lag.max=40, plot=TRUE)
Hydro_pacf=Pacf(ts_RE_data[,3],lag.max=40, plot=TRUE)
```
The values of PACF are smaller than ACFs (except the lag of 1) because the calculation of PACF removes the influence of all these intermediate variables and only leaves the directly correlation between Yt and Yt-h. In all three graphs, there are positive ACFs at some lags with corresponding negative PACFs. 