---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
subtitle: "Assignment 6 - Due date 03/16/22"
author: "Yu Hai"
output: pdf_document
geometry: margin=2.54cm
header-includes:
  - \usepackage{enumerate}
  - \usepackage{enumitem}
---
  
## Directions
  
  You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.  

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A06_Sp22.Rmd"). Submit this pdf using Sakai. 

## Questions 

This assignment has general questions about ARIMA Models. 

Packages needed for this assignment: "forecast","tseries". Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
```


## Q1
Describe the important characteristics of the sample autocorrelation function (ACF) plot and the partial sample autocorrelation function (PACF) plot for the following models: 

\begin{enumerate}[label=(\alph*)]

\item AR(2)

> Answer: For autoregression model with order of 2, the ACF will decary exponentially with time, while the PACF will show a cutoff (huge decrease) at lag=2. 

\item MA(1)

> Answer: For moving average model with order of 1, the ACF will show a cutoff at lag=1, while the PACF will decay exponentially with time. 

\end{enumerate}

## Q2
Recall that the non-seasonal ARIMA is described by three parameters ARIMA$(p,d,q)$ where $p$ is the order of the autoregressive component, $d$ is the number of times the series need to be differenced to obtain stationarity and $q$ is the order of the moving average component. If we don't need to difference the series, we don't need to specify the "I" part and we can use the short version, i.e., the ARMA$(p,q)$. Consider three models: ARMA(1,0), ARMA(0,1) and ARMA(1,1) with parameters $\phi=0.6$ and $\theta= 0.9$. The $\phi$ refers to the AR coefficient and the $\theta$ refers to the MA coefficient. Use R to generate $n=100$ observations from each of these three models

```{r}

ts1 <- arima.sim(list(order = c(1,0,0), ar = 0.6), n = 100)
ts2 <- arima.sim(list(order = c(0,0,1), ma = 0.9), n = 100)
ts3 <- arima.sim(list(order = c(1,0,1), ar = 0.6,ma=0.9), n = 100)
ts.plot(ts1)
```


\begin{enumerate}[label=(\alph*)]

\item Plot the sample ACF for each of these models in one window to facilitate comparison (Hint: use command $par(mfrow=c(1,3))$ that divides the plotting window in three columns).  

```{r}
par(mfrow=c(1,3))
Acf(ts1,main="ARMA(1,0)")
Acf(ts2,main="ARMA(0,1)")
Acf(ts3,main="ARMA(1,1)")
```


\item Plot the sample PACF for each of these models in one window to facilitate comparison.  

```{r}
par(mfrow=c(1,3))
Pacf(ts1,main="ARMA(1,0)")
Pacf(ts2,main="ARMA(0,1)")
Pacf(ts3,main="ARMA(1,1)")
```

\item Look at the ACFs and PACFs. Imagine you had these plots for a data set and you were asked to identify the model, i.e., is it AR, MA or ARMA and the order of each component. Would you be identify them correctly? Explain your answer.

> Answer: Yes. For ARMA (1,0) we can see a graduate decay in ACF and a cutoff at lag=1 in PACF, which measn it is a autoregression model with order=1. For ARMA (0,1), there is a cutoff at lag=1 in ACF while the PACF shows a graduate decay, so it is a moving average model with order=1. For ARMA (1,1), neither the ACF nor the PACF show graduate decay, thus we could identify it is an ARMA model, but we cannot know the order because there is no clear cutoff in ACF or PACF. 

\item Compare the ACF and PACF values R computed with the theoretical values you provided for the coefficients. Do they match? Explain your answer.

> Answer: 


\item Increase number of observations to $n=1000$ and repeat parts (a)-(d).

```{r}
ts4 <- arima.sim(list(order = c(1,0,0), ar = 0.6), n = 1000)
ts5 <- arima.sim(list(order = c(0,0,1), ma = 0.9), n = 1000)
ts6 <- arima.sim(list(order = c(1,0,1), ar = 0.6,ma=0.9), n = 1000)
```

```{r}
#ACF
par(mfrow=c(1,3))
Acf(ts4,main="ARMA(1,0)")
Acf(ts5,main="ARMA(0,1)")
Acf(ts6,main="ARMA(1,1)")
```

```{r}
#PACF
par(mfrow=c(1,3))
Pacf(ts4,main="ARMA(1,0)")
Pacf(ts5,main="ARMA(0,1)")
Pacf(ts6,main="ARMA(1,1)")
```

\end{enumerate}

## Q3

Consider the ARIMA model $y_t=0.7*y_{t-1}-0.25*y_{t-12}+a_t-0.1*a_{t-1}$

\begin{enumerate}[label=(\alph*)]

\item Identify the model using the notation ARIMA$(p,d,q)(P,D,Q)_ s$, i.e., identify the integers $p,d,q,P,D,Q,s$ (if possible) from the equation.

\item Also from the equation what are the values of the parameters, i.e., model coefficients. 
```{r}
#ARIMA(1,1,1)(0,1,1)
#becuase it depends on y(t-1), so p=1, becuase it has seasonality, so d=1, because it depends on a(t-1), so q=1.
#The autoregression coefficient is the coefficient before y(t-1), so it is 0.7, the moving average coefficient is the coefficient before a(t-1), which is -0.1. 
#The seasonal autocorrelation is negative, so Q=1, because P+Q<=1, so P=0. The seasonal component s=12, 
```

\end{enumerate}
## Q4

Plot the ACF and PACF of a seasonal ARIMA$(0, 1)\times(1, 0)_{12}$ model with $\phi =0 .8$ and $\theta = 0.5$ using R. The $12$ after the bracket tells you that $s=12$, i.e., the seasonal lag is 12, suggesting monthly data whose behavior is repeated every 12 months. You can generate as many observations as you like. Note the Integrated part was omitted. It means the series do not need differencing, therefore $d=D=0$. Plot ACF and PACF for the simulated data. Comment if the plots are well representing the model you simulated, i.e., would you be able to identify the order of both non-seasonal and seasonal components from the plots? Explain.

```{r}

```

